Accelerating Precision Medicine with Quantum Computing and AI\\\bigskip \LARGE\mdseries\itshape\color{black!80} Exploring the future of Personalized Health Care\par

### 2022-01-23 paper.tex
(paper) citations + conclusions

% Created 2022-01-23 Sun 14:08
% Intended LaTeX compiler: pdflatex
\documentclass{scrartcl}
\renewcommand\sectionformat{\llap{\thesection\autodot\enskip}}
\renewcommand\subsectionformat{\llap{\thesubsection\autodot\enskip}}
\renewcommand\subsubsectionformat{\llap{\thesubsubsection\autodot\enskip}}

@@

\section{Introduction}
\label{sec:org48ebeae}

Healthcare data—such as information from clinical trials, disease registries, electronic health records (EHRs), and medical devices—is growing at a compound annual growth rate of 36 percent  . Increasingly, this data helps address 4 challenges associated with healthcare:

@@

At the same time, healthcare consumers are making more decisions and have to navigate an increasingly complex system. Traditionally, diagnosing a patient’s condition has been based heavily on the umbrella approach, which is based solely on the patient-reported symptoms, resulting in an umbrella diagnosis and treatment that frequently fails to achieve their intended effects due to individual variability. As a result, many existing therapies fail to achieve their intended effects due to individual variability. Precision medicine aims to allow tailor prevention and treatment with an individual approach. Due to the complexity of human biology, individualized medicine requires taking into account aspects that go well beyond standard medical care. Medical care only has a relative contribution of 10 to 20 percent to outcomes; health-related behaviors, socioeconomic factors, and environmental aspects account for the other 80 to 90 percent. Computationally, the interdependencies and correlations among these diverse contributors create formidable challenges concerning optimizing treatment effectiveness.

Significant investments are being made to deliver the right data and powerful insights at the point of care. Industry incumbents and new entrants alike are trying to create digital experiences that reinforce healthy, preventive behaviors. Despite that, accounting for the exponential possibilities from this diversity of new data is stretching the capabilities of classical computing systems. [\citeprocitem{1}{Grumbling and Horowitz 2020}].

\section{AI in Precision Medicine}
\label{sec:org3e99254}

@@

A key aspect of tailoring medical approaches is proactivity. Early treatments and preventive interventions tend to drastically improve outcomes and optimize costs. Classical machine learning has already shown some promise in predicting the risk of future diseases for a range of patient groups based on EHRs. Supervised and unsupervised quantum-enhanced machine learning techniques could allow earlier more accurate, and more granular risk predictions. Eventually, medical practitioners might even have the tools to understand how an individual’s risk for any given condition changes over time, enabled by continual virtual diagnostics based on ongoing data streams from individuals.

\subsection{Accelerating AI with Quantum Computing}
\label{sec:org9e9021c}

Quantum computers process information in a fundamentally different way from traditional computers. Previous computer technology advancements—such as integrated circuits—enabled faster computing, but were still based on classical information processing. These computers manipulate quantum bits (qubits). These are unlike classical bits, which store information as either a 0 or 1, and they can display uniquely quantum properties, such as entanglement. As a result, it becomes possible to construct quantum algorithms that can outperform their classical counterparts which are not able to leverage quantum phenomena.

\begin{itemize}
\item Chemistry, machine learning/artificial intelligence (AI), optimization, or simulation tasks. Machine learning has shown potential to be enhanced by quantum computing and is symbiotically helping drive quantum advances
\item Complex correlations and interdependencies among many highly interconnected elements, such as molecular structures in which many electrons interact
\item Inherent scaling limits of relevant classical algorithms. For instance, the resource requirements of classical algorithms may increase exponentially with problem size, as is the case when simulating the time evolution of quantum systems.
\end{itemize}

\subsection{Hybrid Quantum-Classical Neural Networks}
\label{sec:org00af154}

@@

\subsection{Generative Adversarial Networks}
\label{sec:org29e917a}

@@

\subsubsection{MolGAN}
\label{sec:org07a4bd5}

@@

\subsubsection{Qubit-efficient Quantum Molecule Generation}
\label{sec:org7497f82}

@@

\section{Prototype}
\label{sec:orgc004852}

@@

\subsection{Image recognition (QuTorch-HG)}
\label{sec:orgb3d0933}

@@

\subsection{Quantum Accelerated Drug \& Molecule Generation (QGAN-HG)}
\label{sec:org00e2ec5}

@@

Firstly, only generated molecules that have high affinity towards the receptor binding sites are considered as valid. Next, a parameterized quantum circuit with last-layer N measures the expectation values and a processes it through a classical stage. Then, we apply the atom layer and bond layer to generate synthetic molecular graphs. Afterwards, a batch of real molecules from a training dataset (in this case we are using QM9) and a batch of generated synthetic molecules are fed into a classical discriminator for real/synthetic prediction and Frechet distance score calculation.

\subsection{Data}
\label{sec:org269d80a}

@@

\subsection{Tools and Hardware}
\label{sec:org2908f15}

@@

While simulated hardware produces \emph{slightly} different results to actual hardware, the margin is within 1\%. The tests below were conducted using actual IBMQ hardware, on the \verb~ibmq_lima~ quantum computer [\citeprocitem{2}{Quantum 2021}]

@@

\section{Results}
\label{sec:orgde3234d}

\subsection{Speed}
\label{sec:orgc8c028e}

@@

\subsection{Accuracy}
\label{sec:org18efb0c}

@@

\subsection{Pricing}
\label{sec:org101ffef}

@@

\section{Conclusion and Further Research}
\label{sec:orgba86c27}

\section{References}
\label{sec:orgc0f0fd3}

\begin{hangparas}{1.5em}{1}
\hypertarget{citeproc_bib_item_1}{\textsc{Grumbling, E. and Horowitz, M.} 2020. \href{https://doi.org/10.17226/25196}{Quantum computing}. .}

\hypertarget{citeproc_bib_item_2}{\textsc{Quantum, I.} 2021. \href{https://quantum-computing.ibm.com/}{Ibm quantum services}. .}
\end{hangparas}

>>>>

% Created 2022-01-23 Sun 15:00
% Intended LaTeX compiler: pdflatex
\documentclass{scrartcl}
\renewcommand\sectionformat{\llap{\thesection\autodot\enskip}}
\renewcommand\subsectionformat{\llap{\thesubsection\autodot\enskip}}
\renewcommand\subsubsectionformat{\llap{\thesubsubsection\autodot\enskip}}

@@

\section{Introduction}
\label{sec:orgd2a88f9}

Healthcare data—such as information from clinical trials, disease registries, electronic health records (EHRs), and medical devices—is growing at a compound annual growth rate of 36 percent (\citeprocitem{9}{NCHIT, 2021}). Increasingly, this data helps address 4 challenges associated with healthcare:

@@

At the same time, healthcare consumers are making more decisions and have to navigate an increasingly complex system. Traditionally, diagnosing a patient’s condition has been based heavily on the umbrella approach, which is based solely on the patient-reported symptoms, resulting in an umbrella diagnosis and treatment that frequently fails to achieve their intended effects due to individual variability. As a result, many existing therapies fail to achieve their intended effects due to individual variability. Precision medicine aims to allow tailor prevention and treatment with an individual approach. (\citeprocitem{11}{Ravizza et al., 2019}). Due to the complexity of human biology, individualized medicine requires taking into account aspects that go well beyond standard medical care. Medical care only has a relative contribution of 10 to 20 percent to outcomes; health-related behaviors, socioeconomic factors, and environmental aspects account for the other 80 to 90 percent (\citeprocitem{6}{Magnan \& , 2017}). Computationally, the interdependencies and correlations among these diverse contributors create formidable challenges concerning optimizing treatment effectiveness. (\citeprocitem{7}{McDermott et al., 2012})

Significant investments are being made to deliver the right data and powerful insights at the point of care. Industry incumbents and new entrants alike are trying to create digital experiences that reinforce healthy, preventive behaviors. Despite that, accounting for the exponential possibilities from this diversity of new data is stretching the capabilities of classical computing systems. (\citeprocitem{3}{Grumbling \& Horowitz, 2020}).

\section{AI in Precision Medicine}
\label{sec:orgcac6fb6}

@@

A key aspect of tailoring medical approaches is proactivity. Early treatments and preventive interventions tend to drastically improve outcomes and optimize costs. Classical machine learning has already shown some promise in predicting the risk of future diseases for a range of patient groups based on EHRs (\citeprocitem{4}{Jauk et al., 2020}). Supervised and unsupervised quantum-enhanced machine learning techniques could allow earlier more accurate, and more granular risk predictions. Eventually, medical practitioners might even have the tools to understand how an individual’s risk for any given condition changes over time, enabled by continual virtual diagnostics based on ongoing data streams from individuals. (\citeprocitem{5}{Kumar et al., 2021})

\subsection{Accelerating AI with Quantum Computing}
\label{sec:org57976b0}

Quantum computers process information in a fundamentally different way from traditional computers. Previous computer technology advancements—such as integrated circuits—enabled faster computing, but were still based on classical information processing. These computers manipulate quantum bits (qubits). These are unlike classical bits, which store information as either a 0 or 1, and they can display uniquely quantum properties, such as entanglement. As a result, it becomes possible to construct quantum algorithms that can outperform their classical counterparts which are not able to leverage quantum phenomena. (\citeprocitem{8}{Montanaro, 2016})

\begin{itemize}
\item Chemistry, machine learning/artificial intelligence (AI), optimization, or simulation tasks. Machine learning has shown potential to be enhanced by quantum computing and is symbiotically helping drive quantum advances (\citeprocitem{2}{Carleo \& Troyer, 2017})
\item Complex correlations and interdependencies among many highly interconnected elements, such as molecular structures in which many electrons interact
\item Inherent scaling limits of relevant classical algorithms. For instance, the resource requirements of classical algorithms may increase exponentially with problem size, as is the case when simulating the time evolution of quantum systems (\citeprocitem{12}{Whitfield et al., 2011}).
\end{itemize}

\subsection{Hybrid Quantum-Classical Neural Networks}
\label{sec:orgaad1700}

@@

\subsection{Generative Adversarial Networks}
\label{sec:org8f70103}

@@

\subsubsection{MolGAN}
\label{sec:orgce38a3d}

@@

\subsubsection{Qubit-efficient Quantum Molecule Generation}
\label{sec:orgac8eb1e}

@@

\section{Prototype}
\label{sec:org2ab3ce7}

@@

\subsection{Image recognition (QuTorch-HG)}
\label{sec:orgeed551a}

@@

\subsection{Quantum Accelerated Drug \& Molecule Generation (QGAN-HG)}
\label{sec:org15b0e26}

@@

Firstly, only generated molecules that have high affinity towards the receptor binding sites are considered as valid. (\citeprocitem{1}{Andreyev \& Zybailov, 2020}) Next, a parameterized quantum circuit with last-layer N measures the expectation values and a processes it through a classical stage. Then, we apply the atom layer and bond layer to generate synthetic molecular graphs. Afterwards, a batch of real molecules from a training dataset (in this case we are using QM9) and a batch of generated synthetic molecules are fed into a classical discriminator for real/synthetic prediction and Frechet distance score calculation.

\subsection{Data}
\label{sec:orgc7a20ea}

@@

\subsection{Tools and Hardware}
\label{sec:org7764740}

@@

While simulated hardware produces \emph{slightly} different results to actual hardware, the margin is within 1\%. The tests below were conducted using actual IBMQ hardware, on the \verb~ibmq_lima~ quantum computer (\citeprocitem{10}{Quantum, 2021})

@@

\section{Results}
\label{sec:orgf7b059b}

\subsection{Speed}
\label{sec:orgd3deb47}

@@

\subsection{Accuracy}
\label{sec:org2aa665c}

@@

\subsection{Pricing}
\label{sec:orgf3f3cb9}

@@

\section{Conclusion and Further Research}
\label{sec:org41345ee}

As you can see Healthcare data, collected from the many different modalities and locations that surround patients—is underused today. Clinical factors only account for a small percentage of health-relevant data. Therefore, obtaining actionable insights from other areas, including genomics, behavior, and social and environmental influences, is critical.

Quantum-enhanced machine learning algorithms stand out for their breadth of application, their efficiency and relatively low running costs. We are entering an era where the characteristics of health datasets, such as being frequently heterogeneous and unevenly distributed, are producing complex computational challenges for modern AI. For instance, researchers have been exploring how to speed up the computationally expensive algorithms that are at the core of machine learning and AI modeling with quantum approaches, including operations with large matrices.

The maturity horizons of quantum computing use cases in healthcare depend not only on cross-industry quantum algorithm and hardware developments, but also on the increasing availability of relevant data. While certain modalities are already being leveraged for advanced computational modeling, health-relevant datasets are rapidly growing from increases in both the number of people being “instrumented,” as well as the data types being collected. As the number of accessible health- relevant data sources continues to grow, the potential for quantum computing to add value likewise increases.

\section{References}
\label{sec:org3c7375e}

\begin{hangparas}{1.5em}{1}
\hypertarget{citeproc_bib_item_1}{Andreyev, D. S., \& Zybailov, B. L. (2020). Integration of flow cytometry and single cell sequencing. \textit{Trends in Biotechnology}, \textit{38}(2), 133–136. \url{https://doi.org/10.1016/j.tibtech.2019.09.002}}

\hypertarget{citeproc_bib_item_2}{Carleo, G., \& Troyer, M. (2017). Solving the quantum many-body problem with artificial neural networks. \textit{Science}, \textit{355}(6325), 602–606. \url{https://doi.org/10.1126/science.aag2302}}

\hypertarget{citeproc_bib_item_3}{Grumbling, E., \& Horowitz, M. (2020). \textit{Quantum computing}. \url{https://doi.org/10.17226/25196}}

\hypertarget{citeproc_bib_item_4}{Jauk, S., Kramer, D., Großauer, B., Rienmüller, S., Avian, A., Berghold, A., Leodolter, W., \& Schulz, S. (2020). Risk prediction of delirium in hospitalized patients using machine learning: An implementation and prospective evaluation study. \textit{Journal of the American Medical Informatics Association}, \textit{27}(9), 1383–1392. \url{https://doi.org/10.1093/jamia/ocaa113}}

\hypertarget{citeproc_bib_item_5}{Kumar, Y., Koul, A., Sisodia, P. S., Shafi, J., Verma, K., Gheisari, M., \& Davoodi, M. B. (2021). Heart failure detection using quantum-enhanced machine learning and traditional machine learning techniques for internet of artificially intelligent medical things. \textit{Wireless Communications and Mobile Computing}, \textit{2021}, 1–16. \url{https://doi.org/10.1155/2021/1616725}}

\hypertarget{citeproc_bib_item_6}{Magnan, S., \& . (2017). Social determinants of health 101 for health care: Five plus five. \textit{NAM Perspectives}, \textit{7}(10). \url{https://doi.org/10.31478/201710c}}

\hypertarget{citeproc_bib_item_7}{McDermott, J. E., Wang, J., Mitchell, H., Webb-Robertson, B.-J., Hafen, R., Ramey, J., \& Rodland, K. D. (2012). Challenges in biomarker discovery: combining expert insights with statistical analysis of complex omics data. \textit{Expert Opinion on Medical Diagnostics}, \textit{7}(1), 37–51. \url{https://doi.org/10.1517/17530059.2012.718329}}

\hypertarget{citeproc_bib_item_8}{Montanaro, A. (2016). Quantum algorithms: an overview. \textit{Npj Quantum Information}, \textit{2}(1). \url{https://doi.org/10.1038/npjqi.2015.23}}

\hypertarget{citeproc_bib_item_9}{NCHIT. (2021). \textit{Non-federal acute care hospital electronic health record adoption}. \url{https://www.healthit.gov/data/quickstats/non-federal-acute-care-hospital-electronic-health-record-adoption}}

\hypertarget{citeproc_bib_item_10}{Quantum, I. (2021). \textit{Ibm quantum services}. \url{https://quantum-computing.ibm.com/}}

\hypertarget{citeproc_bib_item_11}{Ravizza, S., Huschto, T., Adamov, A., Böhm, L., Büsser, A., Flöther, F. F., Hinzmann, R., König, H., McAhren, S. M., Robertson, D. H., Schleyer, T., Schneidinger, B., \& Petrich, W. (2019). Predicting the early risk of chronic kidney disease in patients with diabetes using real-world data. \textit{Nature Medicine}, \textit{25}(1), 57–59. \url{https://doi.org/10.1038/s41591-018-0239-8}}

\hypertarget{citeproc_bib_item_12}{Whitfield, J. D., Biamonte, J., \& Aspuru-Guzik, A. (2011). Simulation of electronic structure hamiltonians using quantum computers. \textit{Molecular Physics}, \textit{109}(5), 735–750. \url{https://doi.org/10.1080/00268976.2011.552441}}
\end{hangparas}
